---
title: "EarthquakeVizR Vignette"
author: "Aziiz Sutrisno"
date: "11/10/2019"
output: rmarkdown::html_vignette
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(# Collapse output blocks
                      collapse = TRUE,
                      comment = "#>",
                      fig.width = 7,
                      fig.height = 7,
                      fig.align = "center",
                      warning = FALSE)
```

This vignette describes what is the `earthquakeVizR` package and how it should be used.

## Description

The `earthquakeVizR` package bundles a set of functions to get, clean, visualize, and analyze the data from the [NOAA's Significant Earthquake dataset](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).

## Usage

### Tidying the data

#### Obtaining the data

The function `get_earthquake_data` allows you to load a snapshot of the earthquakes' data downloaded on September 10, 2017; on the proviso that you may want to download the latest data from the NOAA's Website, you can use the `download_earthquake_data` function (requires internet access).

```{r}

raw_data <- "signif.txt"
```
