---
title: "EarthquakeVizR Vignette"
author: "Aziiz Sutrisno"
date: "`r Sys.Date()`"
output: 
    prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Overview of the EarthquakeVizR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(# Collapse output blocks
                      collapse = TRUE,
                      comment = "#>",
                      fig.width = 7,
                      fig.height = 7,
                      fig.align = "center",
                      warning = FALSE)
```

## Description

The `earthquakeVizR` package is a ollection of functions to cleansed and visualize the data from the [NOAA's Significant Earthquake dataset](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).

## Usage

## Package functions

There are five functions in this package:

- `eq_clean_data()`
- `geom_timeline()`
- `geom_timeline_label()`
- `eq_create_label()`
- `eq_map()`

For the purposes of these examples we will use data from NOAA inside this package.


### Tidying the data

The \code{eq_clean_data} and \code{eq_location_clean} are used to preprocess the data obtained from the __NOAA__ 
Here we use the combination of both functions to produce the tidy dataset

```{r, message=FALSE}
raw_data <- "signif.txt"

# We use this function to fix the date and coordinates
dset <- eq_clean_data(raw_data)

# We use this function to fix the location name
dset_location <- eq_location_clean(dset)

```
__Example Result of the cleansed location, time, and the coordinates__
```{r, echo=FALSE, message=FALSE, warning=FALSE}

dset_ex <- dset_location %>% 
  filter(YEAR == 2010 & !is.na(LOCATION_NAME)) %>%
  head(n = 10) %>% 
  select(LOCATION_NAME, DATE, LATITUDE, LONGITUDE)

kable(dset_ex) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

### Visualize earthquake timeline

We will use the following functions to visualize the earthquake timeline. We use `ggplot2` to construct the functions.
First, we are using `geom_timeline` to plot the `DATE` in X-axis and `COUNTRY` in the Y-axis. Second, we pass on the 
`TOTAL_DEATHS` and `EQ_PRIMARY` as optional color and size.

We then use `geom_timeline_label` to pass on the `LOCATION_NAME`

Here is an example:

```{r fig.width = 7, fig.height = 4, warning=FALSE, message=FALSE}
dset_location %>% filter(COUNTRY == "MEXICO", YEAR > 1990) %>%
     ggplot(aes(x = DATE,
                y = COUNTRY,
                color = as.numeric(TOTAL_DEATHS),
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     geom_timeline_label(aes(label = LOCATION_NAME), n_max = 5) +
     labs(size = "Richter scale", color = "No of deaths")
```


### Using Leaflet to Overlay Earthquake Data on a Map

We use `leaflet` to overlay the earthquakes data on a map. We pass on the dataset to the `eq_map` function, the map is centered around the Country location the we chose. There is an option to pass on the data label by using `eq_create_label()` function. 

```{r fig.width = 9, fig.height = 6}
dset_location %>%  
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```
